Traffic aware Resource Allocation schemes for Multi-Cell \acs{MIMO}-\acs{OFDM} systems

### 2014-10-01 Report/simulation_results.tex
Updated double column
temporary-commit

@@

Table. \ref{tbl-1} outlines the channel seen by the users over each sub-channel followed by the rates assigned by three different algorithms, \ac{Q-WSRME} allocation, \ac{JSFRA} approach and the band-wise \ac{Q-WSRM} scheme using \ac{WMMSE} design \cite{wmmse_shi}. The performance metric used for the comparison is the total number of backlogged bits left over at each slot after the allocation, which is denoted by \me{\chi = \sum_{k = 1}^K \; [ Q_k - t_k ]^+}. Even though \me{\mc{U}(1)} and \me{\mc{U}(3)} has equal number of backlogged packets of \me{Q_{1} = Q_{3} = 4} bits, user \me{\mc{U}(3)} got scheduled in the first sub-channel due to the better channel condition. In contrast to the \ac{Q-WSRME} scheme, the \ac{JSFRA} approach assigns the first user on the first sub-channel, which reduces the total number of backlogged packets waiting at the transmitter. The rate allocated for \me{\mc{U}(2)} on the second sub-channel is higher in \ac{JSFRA} scheme compared to the other schemes. It is due to the efficient allocation of the total power shared across the sub-channels.
\begin{table*}
\centering
\renewcommand{\arraystretch}{1.25} \scriptsize
\begin{tabular}{|*{14}{c|}}
\hline
\multirow{2}{*}{Users} & \multirow{2}{*}{Queued} & \multicolumn{3}{c|}{\multirow{2}{*}{Channel Gains}} & \multicolumn{3}{c|}{Q-WSRME approach} & \multicolumn{3}{c|}{\multirow{2}{*}{JSFRA Scheme}} & \multicolumn{3}{c|}{Q-WSRM band} \\
\multirow{2}{*}{} & \multirow{2}{*}{Packets} & \multicolumn{3}{c|}{} & \multicolumn{3}{c|}{(modified \emph{backpreassure})} & \multicolumn{3}{c|}{} & \multicolumn{3}{c|}{Alloc Scheme} \\
\cline{3-14}
 && SC-\me{1} & SC-\me{2} & SC-\me{3} & SC-\me{1} & SC-\me{2} & SC-\me{3} & SC-\me{1} & SC-\me{2} & SC-\me{3} & SC-\me{1} & SC-\me{2} & SC-\me{3} \\
\hline
\me{1} & \me{4} & \me{1.71} &  \me{0.53}  &  \me{0.56} & \me{0} &  \me{0}  &  \me{0} & \me{4.0} &  \me{0}  &  \me{0} & \me{0} &  \me{0}  &  \me{0} \\
\me{2} & \me{8} & \me{0.39} &  \me{1.41}  &  \me{1.03} & \me{0} &  \me{4.88}  &  \me{3.11} & \me{0} &  \me{5.49}  &  \me{0} & \me{0} &  \me{4.39}  &  \me{3.53} \\
\me{3} & \me{4} & \me{2.34} &  \me{1.26}  &  \me{2.32} & \me{4.0} &  \me{0}  &  \me{0} & \me{0} &  \me{0}  &  \me{4.0} & \me{5.81} &  \me{0}  &  \me{0} \\
\hline
\multicolumn{5}{|c|}{Remaining backlogged packets (\me{\chi})} & \multicolumn{3}{c|}{\me{3.92} bits} & \multicolumn{3}{c|}{\me{2.51} bits} & \multicolumn{3}{c|}{\me{5.89} bits} \\
\hline
\end{tabular}
\caption{Sub channel wise allocation for a scheduling instant}
\label{tbl-1}
\end{table*}
%\begin{figure}
%\centering
%\includegraphics[width=\columnwidth]{tbl-1}
%\caption{Allocation for a \ac{SISO} Model comparing different Algorithms}
%\label{tbl-1}
%\end{figure}

In order to understand the behavior in a \ac{MIMO} framework, we consider a system with \me{N=3} sub-channels and \me{N_B = 3} \acp{BS}, each equipped with \me{N_T = 4} transmit antennas operating at \me{10}dB \ac{SNR}, serving \me{|\mc{U}_b| = 3} users each. The path loss between the \acp{BS} and the users are uniformly generated from \me{[0,-3]} dB and the association is made by selecting the \ac{BS} with lowest path loss component. Fig. \ref{fig-1} shows the performance of the centralized schemes for a single receive antenna system. It compares the total number of \ac{SCA} updates required by the \ac{JSFRA}, \ac{SRA} and the \ac{Q-WSRME} schemes to perform the optimal allocations to minimize the total number of backlogged packets. The total number of queued packets for Fig. \ref{fig-1} is given by \me{Q_k = [14,15,14,8,12,9,12,11,11]} bits and for Fig. \ref{fig-2} is \me{Q_k = [9,12,8,12,5,4,10,8,5]} bits respectively.
\begin{figure*}
\centering
\subfloat[][System \me{\lbrace N,N_B,K,N_T,N_R \rbrace = \lbrace 4,3,9,4,1\rbrace}]{
\includegraphics[width=0.48\textwidth]{fig-1-3-1}
\label{fig-1}}
\hfill
\subfloat[][System \me{\lbrace N,N_B,K,N_T,N_R \rbrace = \lbrace 2,3,9,4,2\rbrace}]{
\includegraphics[width=0.48\textwidth]{fig-2-4-1}
\label{fig-2}}
\caption{Number of backlogged packets at the \ac{SCA} update points}
\label{fig-a}
\end{figure*}

@@

In the scenario defined by Fig. \ref{fig-a}, \ac{Q-WSRME} performs marginally inferior to the \ac{JSFRA} algorithms due to the weights used in the rate maximization algorithm. The performance degradation can be attributed to the fact that the \ac{Q-WSRME} algorithm favors the users with the large number of backlogged packets as compared to the users with better channel conditions. Fig. \ref{fig-2} compares the algorithms for \me{N_R = 2} receive antenna case. In all figures, the receivers are updated along with the \ac{SCA} update instants \textit{i.e}, \me{J_{\max} = 1} in the Algorithm \ref{algo-1}. It is also noted that the performance degradation by performing the group update is very minimal. Since the receiver minimizes the objective for the fixed transmit precoders, the convergence is monotonic as can be seen from the figures.
\begin{table}
\centering
\renewcommand{\arraystretch}{1.25} \scriptsize
\begin{tabular}{|c|*{8}{c}|c|}
\hline
\me{q} & \multicolumn{8}{c|}{user indices} & \me{\chi} \\
\hline
\me{1} & 15.00 & 3.95 & 5.26 & 8.95 & 7.03 & 11.90 & 12.00 & 9.73 & 25.15 \\
\me{2} & 11.23 & 3.93 & 10.76 & 10.65 & 10.27 & 9.68 & 8.77 & 5.90 & 27.77 \\
\me{\infty} & 11.41 & 4.41 & 10.41 & 10.41 & 10.41 & 8.41 &  8.41 &  6.41 & 28.68 \\
\hline
\me{Q_k}  & 15.0 &  8.0 &  14.0 & 14.0 &  14.0 & 12.0 & 12.0 & 10.0  \\
\cline{1-9}
\end{tabular}
\caption{Queue information for the system \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 5,2,8,1 \rbrace}}
\label{tbl-3}
\end{table}
%\begin{figure}
%\centering
%\includegraphics[width=\columnwidth]{tbl-2}
%\caption{Queue information for the system \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 5,2,8,1 \rbrace}}
%\label{tbl-3}
%\end{figure}

>>>>

\begin{table*}
	\centering
	\renewcommand{\arraystretch}{1.25} \scriptsize
	\begin{tabular}{|*{14}{c|}}
		\hline
		\multirow{2}{*}{Users} & \multirow{2}{*}{Queued} & \multicolumn{3}{c|}{\multirow{2}{*}{Channel Gains}} & \multicolumn{3}{c|}{Q-WSRME approach} & \multicolumn{3}{c|}{\multirow{2}{*}{JSFRA Scheme}} & \multicolumn{3}{c|}{Q-WSRM band} \\
		\multirow{2}{*}{} & \multirow{2}{*}{Packets} & \multicolumn{3}{c|}{} & \multicolumn{3}{c|}{(modified \emph{backpreassure})} & \multicolumn{3}{c|}{} & \multicolumn{3}{c|}{Alloc Scheme} \\
		\cline{3-14}
		&& SC-\me{1} & SC-\me{2} & SC-\me{3} & SC-\me{1} & SC-\me{2} & SC-\me{3} & SC-\me{1} & SC-\me{2} & SC-\me{3} & SC-\me{1} & SC-\me{2} & SC-\me{3} \\
		\hline
		\me{1} & \me{4} & \me{1.71} &  \me{0.53}  &  \me{0.56} & \me{0} &  \me{0}  &  \me{0} & \me{4.0} &  \me{0}  &  \me{0} & \me{0} &  \me{0}  &  \me{0} \\
		\me{2} & \me{8} & \me{0.39} &  \me{1.41}  &  \me{1.03} & \me{0} &  \me{4.88}  &  \me{3.11} & \me{0} &  \me{5.49}  &  \me{0} & \me{0} &  \me{4.39}  &  \me{3.53} \\
		\me{3} & \me{4} & \me{2.34} &  \me{1.26}  &  \me{2.32} & \me{4.0} &  \me{0}  &  \me{0} & \me{0} &  \me{0}  &  \me{4.0} & \me{5.81} &  \me{0}  &  \me{0} \\
		\hline
		\multicolumn{5}{|c|}{Remaining backlogged packets (\me{\chi})} & \multicolumn{3}{c|}{\me{3.92} bits} & \multicolumn{3}{c|}{\me{2.51} bits} & \multicolumn{3}{c|}{\me{5.89} bits} \\
		\hline
	\end{tabular}
	\caption{Sub channel wise allocation for a scheduling instant}
	\label{tbl-1}
\end{table*}
%\begin{figure}
%\centering
%\includegraphics[width=\columnwidth]{tbl-1}
%\caption{Allocation for a \ac{SISO} Model comparing different Algorithms}
%\label{tbl-1}
%\end{figure}

@@

Table. \ref{tbl-1} outlines the channel seen by the users over each sub-channel followed by the rates assigned by three different algorithms, \ac{Q-WSRME} allocation, \ac{JSFRA} approach and the band-wise \ac{Q-WSRM} scheme using \ac{WMMSE} design \cite{wmmse_shi}. The performance metric used for the comparison is the total number of backlogged bits left over at each slot after the allocation, which is denoted by \me{\chi = \sum_{k = 1}^K \; [ Q_k - t_k ]^+}. Even though \me{\mc{U}(1)} and \me{\mc{U}(3)} has equal number of backlogged packets of \me{Q_{1} = Q_{3} = 4} bits, user \me{\mc{U}(3)} got scheduled in the first sub-channel due to the better channel condition. In contrast, the \ac{JSFRA} approach assigns the first user on the first sub-channel, which reduces the total number of backlogged packets waiting at the transmitter. The rate allocated for \me{\mc{U}(2)} on the second sub-channel is higher in \ac{JSFRA} scheme compared to the other schemes. It is due to the efficient allocation of the total power shared across the sub-channels.

For the \ac{MIMO} framework, we consider a system with \me{N=3} sub-channels and \me{N_B = 3} \acp{BS}, each equipped with \me{N_T = 4} transmit antennas operating at \me{10}dB \ac{SNR}, serving \me{|\mc{U}_b| = 3} users each. The path loss between the \acp{BS} and the users are uniformly generated from \me{[0,-3]} dB and the association is made by selecting the \ac{BS} with lowest path loss component. Fig. \ref{fig-1} shows the performance of the centralized schemes for a single receive antenna system. It compares the total number of \ac{SCA} updates required by the \ac{JSFRA}, \ac{SRA} and the \ac{Q-WSRME} schemes to perform the optimal allocations to minimize the total number of backlogged packets. The total number of queued packets for Fig. \ref{fig-1} is given by \me{Q_k = [14,15,14,8,12,9,12,11,11]} bits and for Fig. \ref{fig-2} is \me{Q_k = [9,12,8,12,5,4,10,8,5]} bits respectively.
\begin{figure*}
\centering
\subfloat[][System \me{\lbrace N,N_B,K,N_T,N_R \rbrace = \lbrace 4,3,9,4,1\rbrace}]{
\includegraphics[width=0.48\textwidth]{fig-1-4}
\label{fig-1}}
\hfill
\subfloat[][System \me{\lbrace N,N_B,K,N_T,N_R \rbrace = \lbrace 2,3,9,4,2\rbrace}]{
\includegraphics[width=0.48\textwidth]{fig-2-5}
\label{fig-2}}
\caption{Number of backlogged packets at the \ac{SCA} update points}
\label{fig-a}
\end{figure*}

@@

In the scenario defined by Fig. \ref{fig-a}, \ac{Q-WSRME} performs marginally inferior to the \ac{JSFRA} algorithms due to the weights used in the rate maximization algorithm. The performance degradation can be attributed to the fact that the \ac{Q-WSRME} algorithm favors the users with the large number of backlogged packets as compared to the users with better channel conditions. Fig. \ref{fig-2} compares the algorithms for \me{N_R = 2} receive antenna case. In all figures, the receivers are updated along with the \ac{SCA} update instants \textit{i.e}, \me{J_{\max} = 1} in the Algorithm \ref{algo-1}. It is also noted that the performance degradation by performing the group update is very minimal. Since the receiver minimizes the objective for the fixed transmit precoders, the convergence is monotonic as can be seen from the figures.
\begin{table}
\flushleft
\renewcommand{\arraystretch}{1.25} \scriptsize
\begin{tabular}{|c|*{8}{c}|c|}
\hline
\me{q} & \multicolumn{8}{c|}{user indices} & \me{\chi} \\
\hline
\me{1} & 15.0 & 3.95 & 5.26 & 8.95 & 7.0 & 11.9 & 12.0 & 9.7 & 25.15 \\
\me{2} & 11.2 & 3.9 & 10.76 & 10.65 & 10.27 & 9.68 & 8.77 & 5.9 & 27.77 \\
\me{\infty} & 11.4 & 4.4 & 10.4 & 10.4 & 10.4 & 8.4 &  8.4 &  6.4 & 28.68 \\
\hline
\me{Q_k}  & 15.0 &  8.0 &  14.0 & 14.0 &  14.0 & 12.0 & 12.0 & 10.0  \\
\cline{1-9}
\end{tabular}
\caption{Queues for \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 5,2,8,1 \rbrace}}
\label{tbl-3} \vspace{-0.5cm}
\end{table}
%\begin{figure}
%\centering
%\includegraphics[width=\columnwidth]{tbl-2}
%\caption{Queue information for the system \me{\lbrace N,N_B,K,N_R \rbrace = \lbrace 5,2,8,1 \rbrace}}
%\label{tbl-3}
%\end{figure}

### 2014-10-01 Report/sub_channel_wsrm.tex
temporary-commit

The proposed \acl{QM} \ac{SRA} formulation enables us to solve for the transmit precoders of all the users associated with the coordinating \ac{BS} in the set \me{\mathcal{B}} over each sub-channel independently by fixing the transmit power on each sub-channel to a constant value \me{P_{\max,n}} as compared to the global power constraint defined by \eqref{eqn-4.3}. In contrast to the decomposition based approach for the sub-channel wise resource allocation, where the primal/dual variables are exchanged, this method requires the update on the number of queued bits before each sub-channel wise optimization. The total number of queued bits for each user are updated by the difference between the total number of queued bits present during the current slot to the total number of bits that are guaranteed by the earlier sub-channel allocations for the same slot as
\begin{equation}
Q_{k,n} = \max{\Big \lbrace Q_k - \sum_{r = 1}^{n-1} \, \sum_{l = 1}^{L} \, t_{l,k,r} ,0 \Big \rbrace }, \; \forall \; k \in \mathcal{U}
\label{eqn-weight}
\end{equation}
where \me{Q_{k,n}} is the total number of queued bits used in the optimization problem carried out for the sub-channel \me{n}. In the expression \eqref{eqn-weight}, \me{Q_k} denotes the total number of queued bits waiting to be transmitted for the user \me{k} during the current slot and \me{t_{l,k,r}} is the rate or guaranteed bits allocated over the sub-channel \me{r}. However, the proposed scheme is sensitive to the order in which the sub-channels are selected for the optimization problem. The algorithmic representation of the \ac{SRA} scheme is shown in Algorithm \ref{algo-5}
\begin{algorithm}
 \SetAlgoLined
 \DontPrintSemicolon
 \BlankLine
 \SetKwInput{KwInit}{Initialize}
 \KwIn{\me{a_k, \, Q_k, \, \mvec{H}{b,k,n},\; \fall b \in \mathcal{B}, \, \fall k \in \mathcal{U}}}
 \KwIn{permute \me{\mathcal{N} \rightarrow \tilde{\mathcal{N}}}}
 \For{\me{n \leftarrow 1 } \KwTo \me{N}}{
 update \me{Q_{k,n}} using \eqref{eqn-weight} and let \me{\hat{n} = \tilde{\mathcal{N}}(n)}\;
 \KwOut{\me{\mvec{m}{l,k,\hat{n}}} and \me{\mvec{w}{l,k,\hat{n}} \fall l \in \set{1,2,\dotsc,L}}}
 \KwInit{\me{i=0} and \me{\tilde{\mbf{m}}_{l,k,\hat{n}}} randomly satisfying per sub-channel power constraint \me{P_{\max,\hat{n}}}}
 update \me{\mvec{w}{l,k,\hat{n}}} and \me{\tilde{\mbf{u}}_{l,k,\hat{n}}} using \eqref{eqn-10} and \eqref{eqn-8}\;
 \Repeat{Queue convergence or \me{i \geq I_{\max}}}{
 initialize \me{j = 0}\;
 \Repeat{\ac{SCA} convergence or \me{j \geq J_{\max}}}{
 solve for \me{\mvec{m}{l,k,\hat{n}}} using \eqref{eqn-9} with per sub-channel power constraint \me{P_{\max,\hat{n}}}\;
 update the constraint set \eqref{eqn-8} with \me{\tilde{u}_{l,k,\hat{n}}} and \me{\mvec{m}{l,k,\hat{n}}} using \eqref{eqn-wsrm-expr}\;
 $j = j + 1$\;
 }
 update \me{\mvec{w}{l,k,n}} using \eqref{eqn-10} with the updated precoders \me{\mvec{m}{l,k,n}}\;
 $i = i + 1$\;
 }
 }
 \caption{Algorithm of \acs{SRA} scheme}
 \label{algo-5}
\end{algorithm}

>>>>

The proposed \acl{QM} \ac{SRA} formulation enables us to solve for the transmit precoders of all the users associated with the coordinating \ac{BS} in the set \me{\mathcal{B}} over each sub-channel independently by fixing the transmit power on each sub-channel to a constant value \me{P_{\max,n}} as compared to the global power constraint defined by \eqref{eqn-4.3}. In contrast to the decomposition based approach for the sub-channel wise resource allocation, where the primal/dual variables are exchanged, this method requires the update on the number of queued bits before each sub-channel wise optimization. The total number of queued bits for each user are updated by the difference between the total number of queued bits present during the current slot to the total number of bits that are guaranteed by the earlier sub-channel allocations for the same slot as
\begin{equation}
Q_{k,n} = \max{\Big \lbrace Q_k - \sum_{r = 1}^{n-1} \, \sum_{l = 1}^{L} \, t_{l,k,r} ,0 \Big \rbrace }, \; \forall \; k \in \mathcal{U}
\label{eqn-weight}
\end{equation}
where \me{Q_{k,n}} is the total number of queued bits used in the optimization problem carried out for the sub-channel \me{n}. In the expression \eqref{eqn-weight}, \me{Q_k} denotes the total number of queued bits waiting to be transmitted for the user \me{k} during the current slot and \me{t_{l,k,r}} is the rate or guaranteed bits allocated over the sub-channel \me{r}. However, the proposed scheme is sensitive to the order in which the sub-channels are selected for the optimization problem. 
%The algorithmic representation of the \ac{SRA} scheme is shown in Algorithm \ref{algo-5}
%\begin{algorithm}
% \SetAlgoLined
% \DontPrintSemicolon
% \BlankLine
% \SetKwInput{KwInit}{Initialize}
% \KwIn{\me{a_k, \, Q_k, \, \mvec{H}{b,k,n},\; \fall b \in \mathcal{B}, \, \fall k \in \mathcal{U}}}
% \KwIn{permute \me{\mathcal{N} \rightarrow \tilde{\mathcal{N}}}}
% \For{\me{n \leftarrow 1 } \KwTo \me{N}}{
% update \me{Q_{k,n}} using \eqref{eqn-weight} and let \me{\hat{n} = \tilde{\mathcal{N}}(n)}\;
% \KwOut{\me{\mvec{m}{l,k,\hat{n}}} and \me{\mvec{w}{l,k,\hat{n}} \fall l \in \set{1,2,\dotsc,L}}}
% \KwInit{\me{i=0} and \me{\tilde{\mbf{m}}_{l,k,\hat{n}}} randomly satisfying per sub-channel power constraint \me{P_{\max,\hat{n}}}}
% update \me{\mvec{w}{l,k,\hat{n}}} and \me{\tilde{\mbf{u}}_{l,k,\hat{n}}} using \eqref{eqn-10} and \eqref{eqn-8}\;
% \Repeat{Queue convergence or \me{i \geq I_{\max}}}{
% initialize \me{j = 0}\;
% \Repeat{\ac{SCA} convergence or \me{j \geq J_{\max}}}{
% solve for \me{\mvec{m}{l,k,\hat{n}}} using \eqref{eqn-9} with per sub-channel power constraint \me{P_{\max,\hat{n}}}\;
% update the constraint set \eqref{eqn-8} with \me{\tilde{u}_{l,k,\hat{n}}} and \me{\mvec{m}{l,k,\hat{n}}} using \eqref{eqn-wsrm-expr}\;
% $j = j + 1$\;
% }
% update \me{\mvec{w}{l,k,n}} using \eqref{eqn-10} with the updated precoders \me{\mvec{m}{l,k,n}}\;
% $i = i + 1$\;
% }
% }
% \caption{Algorithm of \acs{SRA} scheme}
% \label{algo-5}
%\end{algorithm}
%
